Stats

### 2023-06-06 stats/analysis.Rmd
save model output from stats server

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(viridis) # col scales, plotting...
library(ggpubr)
library(ggthemes)
library(tidyboot) # for bootstrap CIs
library(brms) # modeling
library(tidybayes) # for stat_halfeye

@@

# DEVELOPMENTAL TRAJECTORY
```{r lifespan_sample}
lifespan_sample <- tango_lifespan %>% 
  group_by(age_group) %>%
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = ceiling(sum(female/15)),
  )
```

```{r variation_preparateData}
lifespan_variation <- tango_lifespan %>% 
  group_by(subj_id, age, age_group) %>%
  summarise(mean = mean(click_dist_balloons, na.rm = TRUE))
```

```{r lifespan_plot}
ggplot(data = lifespan_variation, aes(x = age, y = mean)) +
  geom_point(color = "darkgrey", alpha = 0.5) +
  geom_smooth(method = "loess", size = 1, se = F, color = "#006c66") + 
  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 5)) + 
  scale_y_continuous(limits = c(0, 4), breaks = seq(0, 4, 0.5)) + 
  labs(x = "Age in years", y = "Average imprecision in target widths")

ggsave("../figures/tango_lifespan.png", width = 8, height = 4, scale = 1.2)
```

```{r lifespan_model}
mDataLifespan <- tango_lifespan %>% 
  mutate(
    symmetric_position = scale(symmetric_position),
    trial_nr = trial_nr - min(trial_nr), 
    click_dist = abs(click_dist_from_target_center_x),
  ) %>% 
  select(subj_id, age, age_centered, symmetric_position, trial_nr, click_dist, click_dist_balloons)

mLifespanNull <- brm(click_dist_balloons ~ age_centered + symmetric_position + trial_nr + (1 + symmetric_position + trial_nr | subj_id),
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 1000, 
                    iter   = 3000, 
                    chains = 4, 
                    inits  = "random",
                    cores  = 4)

write.csv(mAgeHedge, "../saves/mLifespan.csv", quote = F, row.names = F)
saveRDS(mAgeHedge, "../saves/mLifespan.rds")
```

@@

```{r gazefunnel}
target_position_impr <- tango_model %>% 
  group_by(age_group, target_position) %>% 
  tidyboot_mean(column = click_dist_balloons, na.rm = TRUE)

target_position_impr_ind <- tango_model %>% 
  group_by(subj_id, age_group, target_position) %>% 
  tidyboot_mean(column = click_dist_balloons, na.rm = TRUE)

ggplot() +
  geom_point(data = target_position_impr_ind, aes(x = target_position, y = mean, col = as.factor(age_group)), alpha = .1, position = position_dodge(width = 0.3)) +

  geom_pointrange(data = target_position_impr, aes(x = target_position, y = mean, ymin = ci_lower, ymax = ci_upper, col = as.factor(age_group)), pch = 18,  size = 1, position = position_dodge(width = .3)) +

  scale_y_continuous(limits = c(0, 7), breaks = seq(0, 7, 1)) + 
  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d"), breaks = c("3", "4", "5"), labels = c("3-year-olds", "4-year-olds", "5-year-olds")) +
    labs(x = "Target position (binned)", y = "Average imprecision in target widths", col = "Age in years")

ggsave("../figures/tango_gazefunnel.png", width = 6, height = 4, scale = 1.2)
```

>>>>

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) # data handling etc.
library(viridis) # col scales, plotting...
library(ggpubr)
library(ggthemes)
library(tidyboot) # for bootstrap CIs
library(brms) # modeling
library(tidybayes) # for stat_halfeye
library(cmdstanr)
check_cmdstan_toolchain(fix = TRUE, quiet = TRUE)

@@

# DEVELOPMENTAL TRAJECTORY
```{r lifespan_sample}
lifespan_sample <- tango_lifespan %>% 
  group_by(age_group) %>%
  mutate(female = ifelse(gender == "f", 1, 0)) %>% 
  summarise(
    n_total = n_distinct(subj_id), 
    age_mean = mean(age, na.rm = T) %>% round(2),
    age_sd = sd(age, na.rm = T) %>% round(2),
    age_min = min(age, na.rm = T),
    age_max = max(age, na.rm = T),
    n_females = ceiling(sum(female/15)),
  )

mDataLifespan <- tango_lifespan %>% 
  group_by(subj_id, age, age_centered) %>% 
  summarize(mean_dist = mean(click_dist_balloons, na.rm = TRUE))
```

```{r lifespan_plot}
tango_lifespan %>% 
  group_by(subj_id, age, age_group) %>%
  summarise(mean = mean(click_dist_balloons, na.rm = TRUE)) %>% 
  ggplot(., aes(x = age, y = mean)) +
  geom_point(color = "darkgrey", alpha = 0.5) +
  geom_smooth(method = "loess", size = 1, se = T, color = "#006c66", fill = "#006c66") + 
  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 5)) + 
  scale_y_continuous(limits = c(0, 4), breaks = seq(0, 4, 0.5)) + 
  labs(x = "Age in years", y = "Average imprecision in target widths")

# ggsave("../figures/tango_lifespan.png", width = 8, height = 4, scale = 1.2)
```

```{r lifespan_model}
# model as in tango paper
# except we use click_dist in pixel, lognormal needs > 0
mLifespanLinear <- brm(mean_dist ~ age_centered,
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mLifespanLinear

pp_check(mLifespanLinear)

write.csv(mLifespanLinear, "../saves/mLifespanLinear.csv", quote = F, row.names = F)
saveRDS(mLifespanLinear, "../saves/mLifespanLinear.rds")

## squared age
mLifespanQuadratic <- brm(mean_dist ~ 1 + age_centered + I(age_centered^2),
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mLifespanQuadratic

pp_check(mLifespanQuadratic)

conditional_effects(mLifespanQuadratic, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mLifespanQuadratic, "../saves/mLifespanQuadratic.csv", quote = F, row.names = F)
saveRDS(mLifespanQuadratic, "../saves/mLifespanQuadratic.rds")

## cubic age
mLifespanCubic <- brm(mean_dist ~ 1 + age_centered + I(age_centered^2) + I(age_centered^3),
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mLifespanCubic

pp_check(mLifespanCubic)

plot(mLifespanCubic)

conditional_effects(mLifespanCubic, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mLifespanCubic, "../saves/mLifespanCubic.csv", quote = F, row.names = F)
saveRDS(mLifespanCubic, "../saves/mLifespanCubic.rds")

## gaussian process model
mLifespanGP <- brm(mean_dist ~ gp(age_centered, k=50, c=5/4, scale=TRUE),
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 2000, 
                    iter   = 6000, 
                    chains = 4, 
                    cores  = 4, 
                    threads = threading(8), 
                    control = list(adapt_delta = 0.95), 
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mLifespanGP

pp_check(mLifespanGP)

conditional_effects(mLifespanGP, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mLifespanGP, "../saves/mLifespanGP.csv", quote = F, row.names = F)
saveRDS(mLifespanGP, "../saves/mLifespanGP.rds")

## spline model
# https://discourse.mc-stan.org/t/smooth-spline-modeling-with-brm/6364/2
mLifespanS <- brm(mean_dist ~ s(age_centered),
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 2000,
                    iter   = 6000,
                    chains = 4,
                    cores  = 4,
                    threads = threading(8),
                    control = list(adapt_delta = 0.95),
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

mLifespanS

pp_check(mLifespanS)

conditional_effects(mLifespanS, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

conditional_smooths(mLifespanS2, effects = "weight")

write.csv(mLifespanS, "../saves/mLifespanS.csv", quote = F, row.names = F)
saveRDS(mLifespanS, "../saves/mLifespanS.rds")

## playing around with parameters
mLifespanS2 <- brm(mean_dist ~ s(age_centered, k = 2),
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 2000,
                    iter   = 6000,
                    chains = 4,
                    cores  = 4,
                    threads = threading(8),
                    control = list(adapt_delta = 0.95),
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

conditional_effects(mLifespanS2, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mLifespanS2, "../saves/mLifespanS2.csv", quote = F, row.names = F)
saveRDS(mLifespanS2, "../saves/mLifespanS2.rds")

# 3
mLifespanS3 <- brm(mean_dist ~ s(age_centered, k = 3),
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 2000,
                    iter   = 6000,
                    chains = 4,
                    cores  = 4,
                    threads = threading(8),
                    control = list(adapt_delta = 0.95),
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

conditional_effects(mLifespanS3, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mLifespanS3, "../saves/mLifespanS3.csv", quote = F, row.names = F)
saveRDS(mLifespanS3, "../saves/mLifespanS3.rds")

# 4
mLifespanS4 <- brm(mean_dist ~ s(age_centered, k = 4),
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 2000,
                    iter   = 6000,
                    chains = 4,
                    cores  = 4,
                    threads = threading(8),
                    control = list(adapt_delta = 0.95),
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

conditional_effects(mLifespanS4, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mLifespanS4, "../saves/mLifespanS4.csv", quote = F, row.names = F)
saveRDS(mLifespanS4, "../saves/mLifespanS4.rds")

# 5
mLifespanS5 <- brm(mean_dist ~ s(age_centered, k = 5),
                    data = mDataLifespan,
                    family = lognormal(),
                    warmup = 2000,
                    iter   = 6000,
                    chains = 4,
                    cores  = 4,
                    threads = threading(8),
                    control = list(adapt_delta = 0.95),
                    backend = "cmdstanr") %>% add_criterion(c("loo","waic"))

conditional_effects(mLifespanS5, ndraws = 200, spaghetti = TRUE) %>% 
  plot(points = T)

write.csv(mLifespanS5, "../saves/mLifespanS5.csv", quote = F, row.names = F)
saveRDS(mLifespanS5, "../saves/mLifespanS5.rds")
```

```{r prediction_plot}
# https://ourcodingclub.github.io/tutorials/brms/#plot
library(tidybayes)

mDataLifespan %>%
  add_predicted_draws(mLifespanQuadratic) %>%  # adding the posterior distribution
  ggplot(aes(x = age_centered, y = mean_dist)) +  
  geom_jitter(data = mDataLifespan, colour = "grey", size = 1.5, alpha = 0.8) + 
  stat_lineribbon(aes(y = .prediction), .width = c(.95, .80), alpha = 0.5, colour = "black") +
  scale_fill_brewer(palette = "Greys") + 
  scale_y_continuous(limits = c(0,4))
```

```{r lifespan_modelcomparison}
# mLifespanLinear <- readRDS("../saves/mLifespanLinear.rds")
# mLifespanQuadratic <- readRDS("../saves/mLifespanQuadratic.rds")
# mLifespanCubic <- readRDS("../saves/mLifespanCubic.rds")
# mLifespanGP <- readRDS("../saves/mLifespanGP.rds")
# mLifespanS <- readRDS("../saves/mLifespanS.rds")

mLifespanComparison <- loo_compare(mLifespanLinear, mLifespanQuadratic, mLifespanCubic, mLifespanGP, mLifespanS, mLifespanS2, mLifespanS3, mLifespanS4, mLifespanS5, criterion = "waic") %>% 
  as_tibble(rownames = "model") %>%
  left_join(as_tibble(model_weights(mLifespanLinear, mLifespanQuadratic, mLifespanCubic, mLifespanGP, mLifespanS, mLifespanS2, mLifespanS3, mLifespanS4, mLifespanS5, weights = "waic"), rownames = "model")) %>% 
  rename(modelweight = value)

mLifespanComparison %>% 
  select(-c(se_elpd_waic, p_waic, se_p_waic)) %>% 
  mutate(across(where(is.numeric), round, 4))

write.csv(mLifespanComparison, "../saves/mLifespanComparison.csv", quote = F, row.names = F)
saveRDS(mLifespanComparison, "../saves/mLifespanComparison.rds")
```

```{r lifespan_posteriordraws}
# lifespan_draws <- bind_rows(
#   mLifespanNull %>% gather_draws(b_age_centered, b_trial_nr, b_symmetric_position) %>% mutate(model = "null model"), 
#   # mOrevFull %>% gather_draws(b_age_centered, b_mean_tango_centered) %>% mutate(model = "full model"), 
#   ) %>% 
#   mutate(
#     .variable = factor(.variable, levels = c("b_age_centered", "b_trial_nr", "b_symmetric_position"), labels = c("Age (scaled)", "Trial Nr", "Symmetric Position")), 
#   ) %>% 
#   rename(estimate = .value, predictor = .variable)
```

```{r lifespan_modelplot}
# lifespan_modelplot <- lifespan_draws %>% 
#   ggplot(., aes(x = estimate, y = predictor)) +
#   tidybayes::stat_halfeye(.width = c(.95, .8), scale = 1, slab_color = "#006c66", slab_fill = "#006c66", slab_alpha = 0.4) +
#     labs(x = "Posterior estimate when predicting oREV score", y = "") +
#   guides(fill = "none")+
#   geom_vline(xintercept = 0, linetype = "dashed") +
#   coord_cartesian(expand = TRUE, xlim = c(-0.1, 0.6), ylim = c(1.5, 2.5)) +
#   theme_classic()
# 
# lifespan_modelplot
# 
# saveRDS(lifespan_modelplot, file = "../saves/lifespan_modelplot.rds")
# ggsave("../figures/lifespan_modelplot.png", width = 10, height = 6, scale = 0.7, bg = "white")
```

@@

```{r gazefunnel}
# target_position_impr <- tango_model %>% 
#   group_by(age_group, target_position) %>% 
#   tidyboot_mean(column = click_dist_balloons, na.rm = TRUE) 
# 
# target_position_impr_ind <- tango_model %>% 
#   group_by(subj_id, age_group, target_position) %>% 
#   tidyboot_mean(column = click_dist_balloons, na.rm = TRUE) 
# 
# ggplot() +
#   geom_point(data = target_position_impr_ind, aes(x = target_position, y = mean, col = as.factor(age_group)), alpha = .1, position = position_dodge(width = 0.3)) +
#   
#   geom_pointrange(data = target_position_impr, aes(x = target_position, y = mean, ymin = ci_lower, ymax = ci_upper, col = as.factor(age_group)), pch = 18,  size = 1, position = position_dodge(width = .3)) + 
#   
#   scale_y_continuous(limits = c(0, 7), breaks = seq(0, 7, 1)) + 
#   scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d"), breaks = c("3", "4", "5"), labels = c("3-year-olds", "4-year-olds", "5-year-olds")) +
#     labs(x = "Target position (binned)", y = "Average imprecision in target widths", col = "Age in years")
# 
# ggsave("../figures/tango_gazefunnel.png", width = 6, height = 4, scale = 1.2)
```
